Convective heat transport in stratified atmospheres at low and high Mach number

### 2017-03-21 tex/nusselt_paper.tex
pushes changes made with Ben

%%%%% Body of the paper
\section{Introduction}
\refstepcounter{section}
\label{sec:intro}
Convection transports energy in stellar and planetary atmospheres.
In these objects, flows are compressible and
feel the atmospheric stratification.  While in some systems this stratification 
is negligible, it is significant in regions such as
the convective envelope of the Sun, which spans 14 density scale heights.
In the bulk of these systems, flows are at very low Mach Number (Ma), 
but numerical constraints have restricted most studies of 
compressible convection to high Ma.
These studies have provided important insight into the low temperature, 
high Ma region near the Sun's surface, but few fundamental
properties of low Ma convection which characterize deeper motions
are known.

%Convective heat transport is essential
%in the cores of high mass stars,
%the envelopes of low mass stars, and the atmospheres of terrestrial and 
%jovian planets.  These astrophysical systems have density stratifications
%ranging from a few density scale heights (e.g. massive star cores)
%up to 14 density scale heights in the convective envelopes of low
%mass stars such as the Sun.
%Atmospheric stratification presumably modifies the fundamental
%dynamics and heat transport mechanisms of these atmospheres.
%Understanding the fundamental
%properties of compressible convection in stratified media is essential
%to characterizing systems in astrophysics and planetary sciences.  
%Numerical constraints have often restricted studies of compressible
%convection to moderately high Mach number (Ma), appropriate to regions near 
%the Sun's surface.  Few fundamental properties of
%low Ma convection, which occurs in the deep solar interior, are known.

In the widely-studied \RB (RB) problem of incompressible Boussinesq convection, 
upflows and downflows are symmetrical, and
the conductive flux ($\propto \grad T$) approaches 
zero in the convective interior.
Early numerical experiments of moderate-to-high Ma compressible convection
in two \cite{graham1975, chan&all1982,
hurlburt&all1984, cattaneo&all1990} and three 
\cite{cattaneo&all1991, brummell&all1996} dimensions
revealed that these two hallmark characteristics of RB convection change
significantly when stratification is included.  Downflow lanes
become fast and narrow, and upflow lanes turn into broad, slow upwellings.
Furthermore, the \emph{entropy} gradient is negated by convection 
rather than the temperature gradient, and
a significant conductive flux can exist in the presence of
efficient convection.

In RB convection, there exist two primary dynamical control parameters: 
the Rayleigh number (Ra, the ratio of
buoyant driving to diffusive damping) and the Prandtl number 
(Pr, the ratio of viscous to thermal
diffusivity). These numbers control two useful
measures of turbulence in the evolved solution:
the Reynolds
number (Re, the strength of advection to viscous diffusion)
and the Peclet number (Pe = Re Pr).  Stratified atmospheres
with negative entropy gradients are unstable to convection.
The magnitude of the unstable entropy gradient, or the superadiabaticity,
joins Ra and Pr as an important control parameter.  This 
\emph{superadiabatic excess} \cite{graham1975}, $\epsilon$, which 
sets the scale of the atmospheric entropy gradient,
primarily controls the Ma of the evolved solution.

%In RB convection, there exist two primary control parameters: 
%the Rayleigh number (Ra, the ratio of
%buoyant driving to diffusive damping) and the Prandtl number 
%(Pr, the ratio of viscous to thermal
%diffusivity).  These numbers coupled with the aspect ratio of 
%the physical domain and the boundary conditions
%determine the dynamics of the convection.  In stratified atmospheres, 
%in addition to specifying the equation of state and
%fundamental properties of the gas, the two control parameters of 
%RB convection are joined by the degree of
%stratification across the domain and the characteristic 
%Ma of the convective flows.  
%Polytropically stratified atmospheres, such as those used in 
%early studies \cite{graham1975, chan&all1982, hurlburt&all1984, 
%cattaneo&all1990, cattaneo&all1991, brummell&all1996}, are an ideal extension of
%RB convection into the stratified realm because the two additional 
%control parameters are directly linked to
%basic properties of the atmosphere.  The density stratification is 
%set by the number of density scale heights (\nrho)
%the atmosphere spans, and Ma is controlled 
%by the superadiabatic excess ($\epsilon$),
%the deviation of the polytropic index from the adiabatic polytropic 
%index \cite{graham1975}.

In this letter we study the behavior of convective heat transport, 
quantified by the Nusselt number (Nu), in plane-parallel, 
two- and three-dimensional, polytropically stratified atmospheres.  
We vary $\epsilon$ and Ra while holding Pr, aspect ratio, boundary conditions,
and initial atmospheric stratification
constant.  We also examie the behavior of flow speeds, as quantifed by Ma,
and the level of turbulence in solutions, as quantified by Re.
%We describe experimental methods in section 
%\ref{sec:experiment}, including the construction of atmospheres, 
%equations, and numerical methods.  
%Results are described in section \ref{sec:results} 
%and their implications are discussed
%in section \ref{sec:discussion}.

\section{Experiment} 
\refstepcounter{section}
\label{sec:experiment}
We examine a monatomic ideal gas with an adiabatic index of
$\gamma = 5/3$ whose equation of state is $P = R\rho T$.
This is consistent with the approach used in earlier work
and is the simplest stratified extension of RB.
We study atmospheres which are initially polytropically stratified,
\begin{equation}
\begin{split}
\rho_0(z) &= \rho_{t}(1 + L_z - z)^m, \\
T_0(z)    &= T_{t}(1 + L_z - z),
\label{eqn:polytrope}
\end{split}
\end{equation}
where $m$ is the polytropic index and $L_z$ is the depth of the atmosphere.
The height coordinate, $z$, increases upwards in the range $[0, L_z]$.  We
specify the depth of the atmosphere, $L_z = e^{n_{\rho}/m} - 1$, by choosing
the number of density scale heights, $n_{\rho}$, it spans.
Throughout this letter we set $n_{\rho} = 3$.  Satisfying hydrostatic
equilibrium sets the value of gravity, $g = R T_t (m + 1)$, which is
constant with depth.  We study atmospheres with aspect
ratios of 4 where both the $x$ and $y$ coordinates have the range $[0, 4L_z]$.
In our 2D cases, we only consider $x$ and $z$.

@@

Convective dynamics are controlled by the superadiabaticity
of the atmospheres as well as the atmospheric diffusivities.
The superadiabaticity, or the magnitude of the (negative) 
entropy gradient, is set by 
the superadiabatic excess, $\epsilon = m_{ad} - m$, where 
$m_{ad} = (\gamma - 1)^{-1}$.
The atmospheric thermal diffusivity, $\chi$,
and kinematic viscosity (or viscous diffusivity), $\nu$,
are determined by the Rayleigh number (Ra) and the Prandtl
number (Pr).  We set $\text{Pr } = \nu/\chi = 1$ throughout
this work. The polytropic initial conditions are in
thermal equilibrium, $\kappa_0 \partial_z T_0 =$ const,
so $\kappa_0 = \chi \rho_0$ must be constant. To keep
Pr constant with height, we set $\chi = \chi_t/\rho_0$
and $\nu = \nu_t/\rho_0$.  Under these constraints, the
diffusivity profiles are specified by choosing Ra
at the top of the domain,
\begin{equation}
\text{Ra}_{\text{top}} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu_t\chi_t},
\end{equation}
where $\Delta S_0 = \epsilon\ln (L_z + 1)$ is the entropy difference 
between the top and bottom boundaries and
$c_P = R\gamma(\gamma-1)^{-1}$ is the specific heat 
at constant pressure. The profiles of $\nu$ and $\chi$ are constant with
time, and Ra at the bottom of the atmosphere is greater than Ra$_{\text{top}}$
by a factor of $e^{2n_{\rho}}$.  This formulation leaves
the thermal conductivity, $\kappa = \rho\chi$, and
the dynamic viscosity, $\mu = \rho\chi$, free to evolve
as the density profile changes.  Throughout this letter,
we will use Ra and Ra$_{\text{top}}$ interchangeably.

We decompose our thermodynamic variables such that $T = T_0 + T_1$ and
$\ln\rho = (\ln\rho)_0 + (\ln\rho)_1$, 
and the velocity $\bm{u} = u\hat{x} + v\hat{y} +  w\hat{z}$ 
has no background component.  In our 2D runs, $v = 0$.
We evolve the Fully Compressible Navier-Stokes equations,
\begin{align}
&\begin{aligned}
&\frac{\partial \ln\rho}{\partial t} + \grad\cdot\bm{u} 
    = -\bm{u}\cdot\grad\ln\rho,
	\label{eqn:continuity_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial\bm{u}}{\partial t} + \grad T - 
&\nu\grad\cdot\lilstressT - \lilstressT\cdot\grad\nu = \\
&-\bm{u}\cdot\grad\bm{u} - T\grad\ln\rho + \bm{g} + 
\nu\lilstressT\cdot\grad\ln\rho,
\label{eqn:momentum_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial T}{\partial t} -\frac{1}{c_V}\left(\right.\chi&\left.
    \grad^2 T + \grad T\cdot\grad\chi\right) = \\
	&-\bm{u}\cdot\grad T - (\gamma-1)T\grad\cdot{\bm{u}} \\
	&+ \frac{1}{c_V}\left(\chi\grad T \cdot\grad\ln\rho +
	\nu\left[\lilstressT\cdot\nabla\right]\cdot\bm{u}\right), 
	\label{eqn:energy_eqn}
\end{aligned}
\end{align}
with the viscous stress tensor given by
\begin{equation}
\sigma_{ij} \equiv \left(\frac{\partial u_i}{\partial x_j} + 
\frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij}\grad\cdot\bm{u}\right).
	\label{eqn:stress_tensor}
\end{equation}
Taking an inner product of
(\ref{eqn:momentum_eqn}) with $\bm{u}$ and adding it to 
(\ref{eqn:energy_eqn}) reveals the full energy equation,
\begin{equation}
\frac{\partial}{\partial t}\left(\rho\left[\frac{|\bm{u}|^2}{2} + c_V T + \phi\right]\right) +
\Div{\bm{F}_{\text{conv}} + \bm{F}_{\text{cond}}} = 0,
	\label{eqn:energy_eqn_full}
\end{equation}
where
$
\bm{F}_{\text{conv}} \equiv \bm{F}_{\text{enth}} + \bm{F}_{\text{KE}} + \bm{F}_{\text{PE}} + \bm{F}_{\text{visc}}
$
is the convective flux and $\bm{F}_{\text{cond}} = -\kappa \grad T$
is the conductive flux.
The individual contributions to $\bm{F}_{\text{conv}}$ are the enthalpy flux, 
$\bm{F}_{\text{enth}} \equiv \rho\bm{u}(c_V T + P/\rho)$;
the kinetic energy flux, 
$\bm{F}_{\text{KE}} \equiv \rho|\bm{u}|^2\bm{u}/2$;
the potential energy flux,
$\bm{F}_{\text{PE}} \equiv \rho\bm{u}\phi$ (with $\phi \equiv -gz$);
and the viscous flux, 
$\bm{F}_{\text{visc}} \equiv -\rho\nu\bm{u}\cdot\lilstressT$, and each 
must be considered. 
Understanding how these fluxes interact  
is crucial in characterizing convective heat transport.

We utilize the 
Dedalus\footnote{http://dedalus-project.org/} \cite{burns&all2016} 
pseudospectral framework to time-evolve  
(\ref{eqn:continuity_eqn})-(\ref{eqn:energy_eqn}) 
using an implicit-explicit (IMEX), third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal, periodic) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  
Grid sizes range from
64x256 coefficient points at the lowest values of 
Ra to 1024x4096 coefficient points at Ra $> 10^{7}$ in 2D,
and from 64x128$^2$ to  256x512$^2$ in 3D. 
By using IMEX timestepping, we implicitly step the 
stiff linear acoustic wave contribution and are able to
efficiently study flows at moderate ($\approx 1$) 
and very low ($\approx 10^{-4}$) Ma.  Our equations take the form
of the FC equations in \cite{lecoanet&all2014}, extended to include
$\nu$ and $\chi$ which vary with depth, and we follow the approach there.
This IMEX approach has been successfully 
tested against a nonlinear benchmark  of the compressible 
Kelvin-Helmholtz instability \cite{Lecoanet_et_al_2016_KH}.

We impose impenetrable, stress free, fixed temperature boundary conditions at
the top and bottom of the domain such that 
$w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. 
$T_1$ is initially filled with
random white noise whose magnitude is infinitesimal
compared to $T_0$ and $\epsilon$.
We filter this noise spectrum in coefficient space, 
such that 25\% of the coefficients
have power. We non-dimensionalize our computational domains by setting
all thermodynamic variables to unity at $z = L_z$ by choosing
$R = T_t = \rho_t = 1$.  By this choice, the non-dimensional
grid length scale is the inverse temperature gradient scale and the 
simulation timescale is the isothermal sound crossing time, 
$\tau_I$, of this unit length.
Meaningful convective dynamics occur on 
timescales of the atmospheric buoyancy time,
$t_b = \tau_I \sqrt{L_z/g\epsilon}$. 
All reported results are taken from time averages
over many buoyancy times beginning 100$t_b$ 
after the start of our simulations in order to
assure our results are not biased by the convective transient.

@@

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/ma_v_Ra.png}
\caption{The mean adiabatic Mach number of long-time-averaged profiles
is shown.  Error bars show the full range of Ma over the depth of the
atmosphere.
(a) Ma is plotted as a function of $\epsilon$, the superadiabatic excess,
at various values of Ra.
(b) Ma is plotted as a function of Ra at various values of $\epsilon$.
Larger symbols with white dots designate 3D runs.
\label{fig:ma_v_eps} }
\end{figure}

Solutions were time-evolved until a long time average of the fluxes
showed little
variance with depth. By performing a linear stability analysis, 
we determined that the onset of convection in 2D
occurs at $\text{Ra}_{\text{crit}} = \{11.15, 10.06, 10.97, 10.97\}$ 
for $\epsilon = \{1.0, 0.5, 10^{-4}, 10^{-7}\}$, respectively.  These
onset values hold in 3D.
We studied Rayleigh
numbers $\lesssim \left\{10^7, 10^5, 10^4, 10^3\right\}$Ra$_{\text{crit}}$ 
for $\epsilon = \{0.5, 10^{-4}, 1.0, 10^{-7}\}$, respectively.

We find that the Mach number is a strong function of 
$\epsilon$ and a weak function of Ra.  
In our 2D runs, when Ma $< 1$ we observe a scaling law of
Ma(Ra$, \epsilon) \propto \epsilon^{1/2}$Ra$^{1/4}$, 
but this relation breaks down as the mean
Ma approaches 1 and the flows become supersonic (see Fig. \ref{fig:ma_v_eps}).  
In 3D, Ma appears to be a function of $\epsilon$ alone, with
Ma $\propto \epsilon^{1/2}$.
Furthermore, the value of $\epsilon$ sets the size of the
evolved thermodynamic variables, such that
$T_1/T_0 \propto \epsilon$ and $\rho_1/\rho_0 \propto \epsilon$.  
These fluctuations are very small when $\epsilon$ and Ma are small,
but can be O(1) for values of $\epsilon$ near $m_{\text{ad}}$.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/snapshots_fig.png}
\caption{Characteristic entropy fluctuations in evolved 2D flows roughly
140$t_b$ after the start of simulations. 
The time- and horizontally-averaged profile is removed in all cases.  
(a) An $\epsilon = 10^{-4}$, low Ma flow. (b) An $\epsilon = 0.5$, flow,
in which shock systems can be seen in the upper atmosphere (e.g.
$(x, z) \approx (30, 18-19), (70, 15-19))$.
(c) A high Ra, $\epsilon = 0.5$ case.  Convective structures become
much smaller scale, and shock systems can be see near $(x, z) \approx (30, 16-19),
(65, 17-19)$.
\label{fig:entropy_snapshots} }
\end{figure}

Low Ma flows (e.g. $\epsilon = 10^{-4}$)
display the classic narrow downflow and broad upflow lanes of stratified
convection (Fig. \ref{fig:entropy_snapshots}a).
While it has been suggested that buoyancy breaking as 
a result of pressure forces leads to
the asymmetric nature of up- and downflows
\cite{hurlburt&all1984}, at low $\epsilon$ this 
effect seems to be unimportant, and the convective structures form
as a result of flows obeying mass conservation while traversing
the stratified medium.  
Our choice of fixed-temperature, stress free boundary conditions 
allows the flux at the boundaries to vary, while also allowing
for domain-wide horizontal velocity profiles along the boundaries.
2D runs at $\text{Ra }> 10^3$Ra$_{\text{crit}}$
and $\epsilon = 10^{-4}$ exhibit long-term states of
flux disequilibrium, in which roll solutions such as those pictured in 
Fig. \ref{fig:entropy_snapshots}a
are punctuated by shearing states similar to those previously
reported in two-dimensional RB convection \cite{goluskin&all2014}.  
During both of these states, the system is in flux disequilibrium.
A long time average
over both of these states produces a flat flux profile which
can be sensibly analyzed, but the presence of these time-dependent states
makes it difficult to measure fluid quantities in these higher-Ra
cases at low $\epsilon$.

At high Ma (e.g. $\epsilon = 0.5$), bulk thermodynamic structures are similar but
shock systems form in the upper atmosphere near downflow lanes 
for sufficiently large values of $\epsilon$ and Ra (Fig. \ref{fig:entropy_snapshots}b,c).
Similar shock phenomena were reported in
both two \cite{cattaneo&all1990} and 
three \cite{malagoli&all1990} dimensional polytropic simulations previously.
As Ra is increased to large values 
(Fig. \ref{fig:entropy_snapshots}c), thermodynamic structures 
no longer span the whole domain but rather break up into 
small eddies which traverse the domain multiple
times before diffusing.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/re_and_nu_v_Ra.png}
\caption{
(a) Variation of Nu as Ra increases at high and low $\epsilon$. 
Errors bars indicate the properly normalized range of the
value of Nu as a function of depth and
large error bars indicate a poorly converged solution.
(b) Variation of Re as Ra increases at high and low $\epsilon$.
Re is measured at the midplane of the atmosphere.
Larger symbols with white dots designate 3D runs.
 \label{fig:re_and_nu_v_ra}.
}
\end{figure}

>>>>

%%%%% Body of the paper
\section{Introduction}
\refstepcounter{section}
\label{sec:intro}
Convection transports energy in stellar and planetary atmospheres.
In these objects, flows are compressible and
feel the atmospheric stratification.  While in some systems this stratification 
is negligible, it is significant in regions such as
the convective envelope of the Sun, which spans 14 density scale heights.
In the bulk of these systems, flows are at very low Mach number (Ma), 
but numerical constraints have restricted most studies of 
compressible convection to high Ma.
These prior studies have provided important insight into the low temperature, 
high Ma region near the Sun's surface. Few fundamental
properties of low Ma convection, which characterizes deeper motions,
are known.

In the widely-studied \RB (RB) problem of incompressible Boussinesq convection, 
a negative temperature gradient leads to convective instability.
In the evolved solution, upflows and downflows are symmetrical, the
temperature in the interior becomes isothermal and
the conductive flux ($\propto \grad T$) approaches 
zero there. 
For compressible convection in a stratified atmosphere, a
negative entropy gradient leads to convective instability.
Early numerical experiments of moderate-to-high Ma compressible convection
in two \cite{graham1975, chan&all1982,
hurlburt&all1984, cattaneo&all1990} and three 
\cite{cattaneo&all1991, brummell&all1996} dimensions
revealed that the evolved state differs from the RB case.
Downflow lanes
become fast and narrow, and upflow lanes turn into broad, slow upwellings.
Furthermore, the \emph{entropy} gradient is negated by convection in the interior, and
a significant temperature gradient and conductive flux persist in the presence of
efficient convection.

In RB convection, there exist two primary dynamical control parameters: 
the Rayleigh number (Ra, the ratio of
buoyant driving to diffusive damping) and the Prandtl number 
(Pr, the ratio of viscous to thermal
diffusivity). These numbers control two useful
measures of turbulence in the evolved solution:
the Reynolds
number (Re, the strength of advection to viscous diffusion)
and the Peclet number (Pe, advection vs. thermal diffusion).  
In stratified atmospheres, the magnitude of the unstable entropy gradient, 
joins Ra and Pr as an important control parameter.  This 
\emph{superadiabatic excess} \cite{graham1975}, $\epsilon$,
sets the scale of the atmospheric entropy gradient.
We find here that $\epsilon$ primarily controls the Ma of the evolved solution.

In this letter we study the behavior of convective heat transport, 
quantified by the Nusselt number (Nu), in plane-parallel, 
two- and three-dimensional, polytropically stratified atmospheres.  
We vary $\epsilon$ and Ra while holding Pr, aspect ratio, boundary conditions,
and initial atmospheric stratification
constant.  We also examine the behavior of flow properties, as quantifed by Ma and Re.

\section{Experiment} 
\refstepcounter{section}
\label{sec:experiment}
We examine a monatomic ideal gas with an adiabatic index of
$\gamma = 5/3$ whose equation of state is $P = R\rho T$. In this
work we take $R = 1$.
This is consistent with the approach used in earlier work \cite{graham1975, chan&all1982,
hurlburt&all1984, cattaneo&all1990, cattaneo&all1991, brummell&all1996} 
and is the simplest stratified extension of RB.
We study atmospheres which are initially polytropically stratified,
\begin{equation}
\begin{split}
\rho_0(z) &= \rho_{t}(1 + L_z - z)^m, \\
T_0(z)    &= T_{t}(1 + L_z - z),
\label{eqn:polytrope}
\end{split}
\end{equation}
where $m$ is the polytropic index and $L_z$ is the depth of the atmosphere.
Subscript 0 indicates initial conditions and subscript $t$ indicates values
at $z = L_z$.
The height coordinate, $z$, increases upwards in the range $[0, L_z]$.  We
specify the depth of the atmosphere, $L_z = e^{n_{\rho}/m} - 1$, by choosing
the number of density scale heights, $n_{\rho}$, it spans initially.
Throughout this letter we set $n_{\rho} = 3$.  Satisfying hydrostatic
equilibrium sets the value of gravity, $g = T_t (m + 1)$, which is
constant with depth.  We study atmospheres with aspect
ratios of 4 where both the $x$ and $y$ coordinates have the range $[0, 4L_z]$.
In our 2D cases, we only consider $x$ and $z$.

@@

Convective dynamics are controlled by the superadiabaticity
of the atmospheres as well as the atmospheric diffusivities.
The superadiabaticity, or the magnitude of the (negative) 
entropy gradient, is set by 
the superadiabatic excess, $\epsilon = m_{ad} - m$, where 
$m_{ad} = (\gamma - 1)^{-1}$.
The atmospheric thermal diffusivity, $\chi$,
and kinematic viscosity (viscous diffusivity), $\nu$,
are determined by the Rayleigh number (Ra) and the Prandtl
number (Pr).  We set $\text{Pr } = \nu/\chi = 1$ throughout
this work. The polytropic initial conditions are in
thermal equilibrium, $\kappa_0 \partial_z T_0 =$ const,
so $\kappa_0 = \chi \rho_0$ must be constant. To keep
Pr constant with height, we set $\chi = \chi_t/\rho_0$
and $\nu = \nu_t/\rho_0$.  Under these constraints, the
diffusivity profiles are specified by choosing Ra
at the top of the domain,
\begin{equation}
\text{Ra}_{\text{top}} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu_t\chi_t},
\end{equation}
where $\Delta S_0 = \epsilon\ln (1 + L_z)$ is the initial entropy difference 
between the top and bottom of the atmosphere and
$c_P = \gamma(\gamma-1)^{-1}$ is the specific heat 
at constant pressure. The profiles of $\nu$ and $\chi$ are constant with
time, and Ra at the bottom of the atmosphere is greater than Ra$_{\text{top}}$
by a factor of $e^{2n_{\rho}}$.  This formulation leaves
the thermal conductivity, $\kappa = \rho\chi$, and
the dynamic viscosity, $\mu = \rho\nu$, free to evolve
as the density profile changes.  Henceforth,
when we use Ra, we are referring to Ra$_{\text{top}}$.

We evolve the Fully Compressible Navier-Stokes equations,
\begin{align}
&\begin{aligned}
&\frac{\partial \ln\rho}{\partial t} + \grad\cdot\bm{u} 
    = -\bm{u}\cdot\grad\ln\rho,
	\label{eqn:continuity_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial\bm{u}}{\partial t} + \grad T - 
&\nu\grad\cdot\lilstressT - \lilstressT\cdot\grad\nu = \\
&-\bm{u}\cdot\grad\bm{u} - T\grad\ln\rho + \bm{g} + 
\nu\lilstressT\cdot\grad\ln\rho,
\label{eqn:momentum_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial T}{\partial t} -\frac{1}{c_V}\left(\right.\chi&\left.
    \grad^2 T + \grad T\cdot\grad\chi\right) = \\
	&-\bm{u}\cdot\grad T - (\gamma-1)T\grad\cdot{\bm{u}} \\
	&+ \frac{1}{c_V}\left(\chi\grad T \cdot\grad\ln\rho +
	\nu\left[\lilstressT\cdot\nabla\right]\cdot\bm{u}\right), 
	\label{eqn:energy_eqn}
\end{aligned}
\end{align}
with the viscous stress tensor given by
\begin{equation}
\sigma_{ij} \equiv \left(\frac{\partial u_i}{\partial x_j} + 
\frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij}\grad\cdot\bm{u}\right).
	\label{eqn:stress_tensor}
\end{equation}
Taking an inner product of
(\ref{eqn:momentum_eqn}) with $\rho\bm{u}$ and adding it to 
$\rho c_V\times$(\ref{eqn:energy_eqn}) reveals the full energy equation,
\begin{equation}
\frac{\partial}{\partial t}\left(\rho\left[\frac{|\bm{u}|^2}{2} + c_V T + \phi\right]\right) +
\Div{\bm{F}_{\text{conv}} + \bm{F}_{\text{cond}}} = 0,
	\label{eqn:energy_eqn_full}
\end{equation}
where
$
\bm{F}_{\text{conv}} \equiv \bm{F}_{\text{enth}} + \bm{F}_{\text{KE}} + \bm{F}_{\text{PE}} + \bm{F}_{\text{visc}}
$
is the convective flux and $\bm{F}_{\text{cond}} = -\kappa \grad T$
is the conductive flux.
The individual contributions to $\bm{F}_{\text{conv}}$ are the enthalpy flux, 
$\bm{F}_{\text{enth}} \equiv \rho\bm{u}(c_V T + P/\rho)$;
the kinetic energy flux, 
$\bm{F}_{\text{KE}} \equiv \rho|\bm{u}|^2\bm{u}/2$;
the potential energy flux,
$\bm{F}_{\text{PE}} \equiv \rho\bm{u}\phi$ (with $\phi \equiv -gz$);
and the viscous flux, 
$\bm{F}_{\text{visc}} \equiv -\rho\nu\bm{u}\cdot\lilstressT$.
Understanding how each of these fluxes interact  
is crucial in characterizing convective heat transport.

We utilize the 
Dedalus\footnote{http://dedalus-project.org/} \cite{burns&all2016} 
pseudospectral framework to time-evolve  
(\ref{eqn:continuity_eqn})-(\ref{eqn:energy_eqn}) 
using an implicit-explicit (IMEX), third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
We decompose our thermodynamic variables such that $T = T_0 + T_1$ and
$\ln\rho = (\ln\rho)_0 + (\ln\rho)_1$, 
and the velocity $\bm{u} = u\hat{x} + v\hat{y} +  w\hat{z}$.
In our 2D runs, $v = 0$.
Subscript 0 variables, set by (\ref{eqn:polytrope}), 
have no time derivative and vary only in $z$.
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal, periodic) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  
Domain sizes range from
64x256 coefficients at the lowest values of 
Ra to 1024x4096 coefficients at Ra $> 10^{7}$ in 2D,
and from 64x128$^2$ to  256x512$^2$ in 3D. 
By using IMEX timestepping, we implicitly step the 
stiff linear acoustic wave contribution and are able to
efficiently study flows at high ($\sim 1$) 
and low ($\sim 10^{-4}$) Ma.  Our equations take the form
of the FC equations in \cite{lecoanet&all2014}, extended to include
$\nu$ and $\chi$ which vary with depth, and we follow the approach there.
This IMEX approach has been successfully 
tested against a nonlinear benchmark  of the compressible 
Kelvin-Helmholtz instability \cite{Lecoanet_et_al_2016_KH}.

We impose impenetrable, stress free, fixed temperature boundary conditions at
the top and bottom of the domain such that 
$w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. 
$T_1$ is initially filled with
random white noise whose magnitude is infinitesimal
compared to $T_0$ and $\epsilon$.
We filter this noise spectrum in coefficient space, 
such that only the lower 25\% of the coefficients
have power. We non-dimensionalize our computational domains by setting
all thermodynamic variables to unity at $z = L_z$ choosing
$R = T_t = \rho_t = 1$.  By this choice, the non-dimensional
length scale is the inverse temperature gradient scale and the 
timescale is the isothermal sound crossing time, 
$\tau_I$, of this unit length.
Meaningful convective dynamics occur on 
timescales of the atmospheric buoyancy time,
$t_b = \tau_I \sqrt{L_z/g\epsilon}$. 
All reported results are taken from time averages
over many $t_b$ beginning \{100, 40\}$t_b$
after the start of our \{2D, 3D\} simulations in order to
assure our results are not biased by the convective transient.

@@

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/ma_v_Ra.png}
\caption{The mean adiabatic Mach number of long-time-averaged profiles
is shown.  Error bars show the full range of Ma over the depth of the
atmosphere.
(a) Ma, at various values of Ra, 
is plotted as a function of $\epsilon$.
(b) Ma, at various values of $\epsilon$, 
is plotted as a function of Ra/Ra$_{\text{crit}}$.
Larger symbols with white dots designate 3D runs.
\label{fig:ma_v_eps} }
\end{figure}

Solutions were time-evolved until a long time average of the fluxes
showed little
variance with depth. By performing a linear stability analysis, 
we determined that the onset of convection
occurs at $\text{Ra}_{\text{crit}} = \{11.15, 10.06, 10.97, 10.97\}$ 
for $\epsilon = \{1.0, 0.5, 10^{-4}, 10^{-7}\}$, respectively.

We measure the adiabatic Mach number (Ma = $|\bm{u}|/\sqrt{\gamma T}$).
We find that Ma is a strong function of 
$\epsilon$ and a weak function of Ra.  
In our 2D runs, when Ma $< 1$ we observe a scaling law of
Ma(Ra$, \epsilon) \propto \epsilon^{1/2}$Ra$^{1/4}$.
this relation breaks down as the mean
Ma approaches 1 (see Fig. \ref{fig:ma_v_eps}).  This transition
occurs near Ra/Ra$_\text{crit} \approx \{10^{2}, 10^{3}\}$ for $\epsilon = \{1, 0.5\}$.
In our limited 3D runs, Ma appears to be a function of $\epsilon$ alone, with
Ma $\propto \epsilon^{1/2}$, so at high Ra, Ma$_{\text{3D}} < \text{Ma}_{\text{2D}}$.
The value of $\epsilon$ also sets the size of the
evolved thermodynamic fluctuations, such that
$T_1/T_0 \propto \epsilon$ and $\rho_1/\rho_0 \propto \epsilon$.  
These fluctuations are very small when $\epsilon$ and Ma are small,
but can be O(1) for values of $\epsilon$ near 1.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/snapshots_fig.png}
\caption{Characteristic entropy fluctuations in evolved 2D flows roughly
140$t_b$ after the start of simulations. 
The time- and horizontally-averaged profile is removed in all cases.  
(a) A low Ma flow at moderate Ra. (b) A high Ma flow at the same Ra as in (a).
(c) a high Ma flow at high Ra.
Shock systems can be seen in the upper atmosphere of the high Ma flows,
for example at $(x, z) \approx (70, 15-19)$ in (b) and $(x, z) \approx 
(65, 17-19)$ in (c).
\label{fig:entropy_snapshots} }
\end{figure}

Low Ma flows (e.g., $\epsilon = 10^{-4}$)
display the classic narrow downflow and broad upflow lanes of stratified
convection (Fig. \ref{fig:entropy_snapshots}a).
At high Ma (e.g., $\epsilon = 0.5$), bulk thermodynamic structures are similar but
shock systems form in the upper atmosphere near downflow lanes 
for sufficiently large values of $\epsilon$ and Ra (Fig. \ref{fig:entropy_snapshots}b,c).
Similar shock phenomena were reported in
both two \cite{cattaneo&all1990} and 
three \cite{malagoli&all1990} dimensional polytropic simulations previously.
As Ra is increased to large values 
(Fig. \ref{fig:entropy_snapshots}c), thermodynamic structures 
no longer span the whole domain but rather break up into 
small eddies which traverse the domain multiple
times before diffusing.  Our choice of fixed-temperature, stress free boundary conditions 
allows the flux at the boundaries to vary, while also allowing
for mean horizontal velocity profiles along the boundaries (winds).
We find these windy convective states, reminiscent of similar states
in RB convection \cite{goluskin&all2014}, at $\epsilon = 10^{-4}$ for high
values of Ra. Roll states (Fig. \ref{fig:entropy_snapshots}a) with high convective
transport alternate with windy states of lower transport.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/re_and_nu_v_Ra.png}
\caption{
Flow properties at high and low $\epsilon$. 
(a) Nu vs. Ra/Ra$_{\text{crit}}$.
Errors bars indicate the variance of Nu with depth;
large error bars indicate a poorly converged solution.
(b) Re vs. Ra/Ra$_{\text{crit}}$.
Re is measured at the midplane of the atmosphere.
Larger symbols with white dots designate 3D runs.
 \label{fig:re_and_nu_v_ra}
}
\end{figure}
